---
title: "Combating Auto Thefts"
author: "Shiyuan Zhou, Xinpeng Shan, Xinyi Chen, Tala ELkukhun, TUT0206, Group3"
subtitle: Investigating attritubes of areas prone to auto thefts
output:
  ioslides_presentation: default
  beamer_presentation: default
  widescreen: yes
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
auto_thefts <- read_csv("auto_thefts.csv")
neighbourhood_profiles_2016 <- read_csv("neighbourhood_profiles_2016.csv")
crime_rates <- read_csv("Neighbourhood_Crime_Rates_Boundary_File_final1.csv")
month_conclusion <- read_csv("month_conclusionfinal1.CSV")
```

## Introduction

- Given two data sets on auto thefts in numerous neighbourhoods in Toronto for over 11,000 vehicles between 2014 and 2018 

- Thefts of vehicles are a major concern 

- Hoping to analyze the public data critically and identify both trends and anomalies

- Suggest functional ways on how to prevent auto theft


## Objectives 

- Is there a significant difference between the number of auto theft occurrences for each month
 from 2014-2018?
- What characteristics do the top 20 neighbourhoods with the highest number of auto theft occurrences have in common?

## Data Summary (Data Wrangling)

- Joined the auto_thefts dataset, the neighbourhood_profiles_2016 dataset, and the table we created with the number of occurrences in each neighbourhood

- Imported another public dataset from TPS that has information about the number of cases for all crime types in neighbourhoods

- Joined the imported dataset with our existing joint data set

---

### **We created three new variables: **
- number of auto theft occurrences per neighbourhood

- occurrence category for the number of occurrences (ex. high) 

- total number of other theft cases over 5 years

```{r, echo=FALSE}
#creating data set that calculates the total number of auto thefts
count_HOODID_differences<-auto_thefts %>%
  group_by(Hood_ID) %>%
  summarise(num_occurrences=n()) %>% arrange (desc(num_occurrences))

#create a joint file that has all info from both data sets, including total number of auto thefts
both_data<-inner_join(x=auto_thefts, y=neighbourhood_profiles_2016, by="Hood_ID")
all_data<-inner_join(x=both_data, y=count_HOODID_differences, by="Hood_ID")

#create a variable that categorizes the neighbourhoods between top 20 most prone to auto thefts, and the rest
#the number 211 is from the data set count_HOODID_differences, it is the num of auto thefts for the 20th neighbourhood in ranking
all_data <- all_data %>%
  mutate(occurrence_cat = ifelse(num_occurrences >= 211, 'Top 20 Neighbourhoods with the Most Number of Auto Thefts', 'Other Neighbourhoods'))
```

## What characteristics do the top 20 theft prone neighbourhoods have?
```{r, echo=FALSE, fig.height=4.5, fig.width=7}
ggplot(data = all_data) +
  aes(x = Lat, y = Long, colour = occurrence_cat) +
  geom_point() +
  labs(x= "Latitude", y="Longitude", colour = "Case Occurred in...",
  title = "Lat-Long Map Illustrating Cases in Different Neighbourhoods")
```


## **Relationship between police division and auto theft occurrences**
```{r,echo=FALSE}
ggplot(data = all_data) +
aes(x = Division ) +
geom_bar(color="black", fill="gray") + 
labs(x= "Police Division", y="Number of auto theft occurrences")
  
```


---

### **When does autotheft happen most frequently?**
- Summary table showing number of autotheft per month from 2014 to 2018

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
auto_thefts3 <- auto_thefts %>%
  filter(!is.na(occurrencemonth))
  
auto_thefts2 <- auto_thefts %>%
  select(occurrencemonth) %>%
  group_by(occurrencemonth)%>%
  summarise(number_autotheft = sum(occurrencemonth == occurrencemonth))%>%
  filter(!is.na(number_autotheft))%>%
  arrange(desc(number_autotheft))
auto_thefts2

```

---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5}
ggplot(data = auto_thefts3,
       aes(x = occurrencemonth)) +
  geom_bar() + coord_flip() + labs(title = "Number of auto thefts per month from 2014 to 2018", y = "number of auto thefts") + theme(text=element_text(size=15)) 
```


## Statistical Methods

- Ran a hypothesis test to determine if there is a significant difference between the number of occurrences of each month over all five years

- Created a linear regression between pop density per square km.           
  Produced the $R^2$ value that provides information on how well our model captured the variability in the data


## Hypothesis Testing

- Null hypothesis($H_0$):$P_m$ = 1/12
- Alternative hypothesis($H_A$):$P_m$ $\neq$ 1/12                           where $P_m$ is the proportion of crime occurence in each month
  
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
n_observations <- 18175 # number of obeservations
repetitions <- 5000 # 1000 simulations
simulated_stats <- rep(NA, repetitions) # 1000 missing values
set.seed(13)
for (i in 1:repetitions){
new_sim <- sample(c("1month", "othermonth"),
                    size = n_observations,
                    prob = c(1/12,11/12),
                    replace = TRUE)
  sim_p <- sum(new_sim == "1month") / n_observations
  simulated_stats[i] <- sim_p; # add new value to vector of results
}

sim <- data_frame(p_1month = simulated_stats)

sim %>% ggplot(aes(x = p_1month)) +
  geom_histogram(colour = "black", fill = "grey") +
  xlab("Simulated proportions of autotheft in one month
       if p=1/12 (samples of size n=18175)")
```

## p-value of 12 months ranked according to amount of auto theft ocurrence form highest month to lowest

```{r, echo=FALSE}
pvalue_11 <- sim %>%
  filter(p_1month >= 0.09607 | p_1month <= 0.070596) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_11)


pvalue_10 <- sim %>%
  filter(p_1month >= 0.09601 | p_1month <= 0.07065) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_10)


pvalue_8 <- sim %>%
  filter(p_1month >= 0.09238 | p_1month <= 0.07429) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_8)


pvalue_7 <- sim %>%
  filter(p_1month >= 0.08941| p_1month <= 0.07726 ) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_7)


pvalue_9 <- sim %>%
  filter(p_1month >= 0.085997| p_1month <= 0.080669 ) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_9)


pvalue_6 <- sim %>%
  filter(p_1month >= 0.08440165 | p_1month <= 0.08226468) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_6)

```
---
```{r, echo=FALSE}
pvalue_12 <- sim %>%
  filter(p_1month >= 0.08380553 | p_1month <= 0.08286107) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_12)


pvalue_5 <- sim %>%
  filter(p_1month >= 0.08595136 | p_1month <= 0.08071527) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_5)


pvalue_3 <- sim %>%
  filter(p_1month >= 0.0864463 | p_1month <= 0.08022) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_3)


pvalue_4 <- sim %>%
  filter(p_1month >= 0.0902433 | p_1month <= 0.076432) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_4)


pvalue_1 <- sim %>%
  filter(p_1month >= 0.0965 | p_1month <= 0.0701) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_1)


pvalue_2 <- sim %>%
  filter(p_1month >=0.1013| p_1month <= 0.0653) %>%
  summarise(p_value = n() / repetitions)
as.numeric(pvalue_2)

```

## Linear Regression between Population Density and Number of Auto Thefts

### Estimating the Coeffcients of the Model
```{r, echo=FALSE}
mod_popdensity <- lm(num_occurrences ~ pop_density_per_square_km, data = all_data)
summary(mod_popdensity)$coefficients
```
---

### Adding the fitted line to the plot...
```{r, echo=FALSE, fig.height= 4}
all_data %>% ggplot(aes(x=pop_density_per_square_km, y=num_occurrences)) + geom_point() +
  geom_smooth(method="lm", se=FALSE) + theme_minimal() +
  labs(x= "Population density per square km", y="Total number of overall thefts",
  title = "Relationship between population density and number of auto theft cases")
```

### How accurate is our model? ($R^2$)
```{r, echo=FALSE}
summary(mod_popdensity)$r.squared
```

## Results

### **Is there a significant difference of auto thefts in each month?**
```{r, echo=FALSE}
month_conclusion
```

---

```{r,echo=FALSE, warning = FALSE, fig.height= 2.5, fig.width= 4}
num_occurrences_above_211 <- all_data %>% 
filter(num_occurrences>=211)
#boxplot for comparision of distribution of population density between the top 20 neighbourhoods with highest auto theft cases and all cases
num_occurrences_above_211 %>% ggplot(aes(x=pop_density_per_square_km, y=num_occurrences)) +
  geom_boxplot() + labs(x= "Pop-density (top 20)", y = "number of auto thefts")

all_data %>% ggplot(aes(x=pop_density_per_square_km,y=num_occurrences)) +geom_boxplot() + labs(x= "Pop-density (all neighbourhood)", y = "number of auto thefts")
```

```{r,echo=FALSE, warning = FALSE, fig.height= 2.5, fig.width= 4}
#boxplot for comparision of distribution of num of private swellings between the top 20 neighbourhoods with highest auto theft cases and all cases
num_occurrences_above_211 %>% ggplot(aes(x=num_private_dwellings, y=num_occurrences)) +
  geom_boxplot() + labs(x= "Private dwellings (top 20)",y = "number of auto thefts")

all_data %>% ggplot(aes(x=num_private_dwellings,y=num_occurrences)) +geom_boxplot()+ labs(x= "Private dwellings (all neighbourhood)", y = "number of auto thefts")
```

## Which premise type has the most amount of auto thefts?
```{r,echo=FALSE}
auto_thefts %>%
ggplot(aes(x=premisetype)) + geom_bar() + labs(y = "number of auto thefts")
```

## What about the other type of thefts?
```{r, echo=FALSE}
#create a variable that sums the total number of theft overall cases for each neighbourhood
crime_rates <- crime_rates %>%
  filter(!is.na (Theft_Over_2014)) %>%
  filter(!is.na (Theft_Over_2015)) %>%
  filter(!is.na (Theft_Over_2016)) %>%
  filter(!is.na (Theft_Over_2017)) %>%
  filter(!is.na (Theft_Over_2018)) %>%
  mutate(total_num_other_theft = Theft_Over_2014 + Theft_Over_2015 + Theft_Over_2016 + Theft_Over_2017+ Theft_Over_2018)
```

```{r, echo=FALSE}
#join two files
crime_auto_file <-inner_join(x=crime_rates, y= all_data, by="Hood_ID")
```

```{r, echo=FALSE, fig.height=5, fig.width=8}
ggplot(data = crime_auto_file,) +
  aes(x = num_occurrences, y = total_num_other_theft, colour = occurrence_cat) +
  geom_point() +
  labs(x= "Total number of auto thefts", y="Total number of overall thefts", colour = "Neighbourhood Category",
  title = "Relationship between number of auto thefts and number of overall thefts\nper neighbourhood from 2014-2018")
```

## Neighbourhoods with highest number of other thefts 

```{r, echo =FALSE}
#table comparing total number of theft overall cases and neighbourhood
theft_over_rank <- crime_rates %>% 
  select(total_num_other_theft, Neighbourhood, Hood_ID) %>%
  arrange (desc(total_num_other_theft))

head(theft_over_rank)

```


## Neighbourhoods with highest number of auto thefts

```{r, echo =FALSE}
#table comparing total number of auto theft and neighbourhood
auto_theft_rank <- count_HOODID_differences %>% 
  select(num_occurrences, Hood_ID) %>%
  arrange (desc(num_occurrences))

head(auto_theft_rank, 11)

#the above two tables are used to get the two neighbourhoods with high auto thefts and high theft overalls
```

## Conclusion

- We could **increase** police force for November, October and August and direct some of police forces on other crimes for July, April, February and January.

- We also discovered that the premise type with the most auto theft occurrences was **outside**, the TPS could establish some concentrated parking areas in order to allow for supervision for outside parks and streets. 

- We thought that more populated neighbourhoods would have a higher number of occurrences but the **opposite was revealed**, so the TPS should be aware that areas with more people do not necessarily mean that they are prone to auto thefts. 

---

- We found that the top 20 neighbourhoods with the highest number of auto thefts had a **higher average population density** and a **higher average number of private dwellings** compared to those of all neighbourhoods, the TPS could notify residence areas in these areas to be cautious and promote the usage of digital video recorder.

- **Black creek and Willowridge-Martingrove-Richview** which were ranked in the top 20 of all neighbourhoods for highest auto theft occurrences and also had a high rate of other thefts, the TPS could focus on reinforcing and educating the areas about theft in general. 

- **limitation:** Most of our answers were based off of the top 20 neighbourhoods with the highest number of auto thefts. This is a limitation as we did not account for the average number of auto thefts for all years. Therefore, all of our conclusions are not always beyond the doubt.


